A* with Perfect Potentials

### 2020-04-20 paper/ch_potentials.tex
A* consistent text improved
restored section about deg 1

A* star is a well-known extension to Dijkstra's algorithm.
It uses a potential function $\pi:V\rightarrow \mathbb{R}^+$.
The potential is often called heuristic.
A* is optimal, if the potential $\pi(v)$ is consistent \cite{Pearl, Judea (1984). Heuristics: Intelligent Search Strategies for Computer Problem Solving. Addison-Wesley. ISBN 0-201-05594-5.}, which essentially means that $\pi(v)$ is a lower bound of the distance from $v$ to $t$.

@@

>>>>

A* star is a well-known extension to Dijkstra's algorithm.
It uses a potential function $\pi:V\rightarrow \mathbb{R}^+$.
The potential is often called heuristic.
A* is optimal, if the potential $\pi(v)$ is consistent \cite{Pearl, Judea (1984). Heuristics: Intelligent Search Strategies for Computer Problem Solving. Addison-Wesley. ISBN 0-201-05594-5.}.
In this paper, all considered potentials are derived from exact shortest path distances and are thus consistent.

@@

\subsection{Avoid Dead-Ends using Largest Biconnected Component}

A lot of nodes in road networks lead to dead-ends.
Unless the source or target is is in a dead-end, it is unnecessary to explore these nodes.
This reduces the number of explored nodes and thus fewer potentials are evaluated.

In the preprocessing phase, before the CH is created, we compute the subgraph $G_C$, called \emph{core}, induced by the largested biconnected component of the undirected graph underlying $G$.
We do this using Tarjan's algorithm \cite{Tarjan's algorithm}.
For every node $v$ in the input graph $G$, we store the attachment node $a_v$ to the core.
For nodes in the core, $a_v=v$.
We exploit that all attachment nodes are single node separators and the problem can be decomposed along them.
The basic CH-potential preprocessing step is only executed for $G_C$.

The query phase is divided into four steps.
In the first step, we use Dijkstra's algorithm to explore the component that contains $s$ until $a_s$ is reached.
If this search finds $t$, then $s$ and $t$ are part of the same component.
An optimal path was found and therefore the algorithm stops early.
Otherwise, in the second step starts.
It consits of applying the basic CH-potential algorithm restricted to $G_C$ from $a_s$ to $a_t$.
Finally, in the third step, we search a path from $a_t$ towards $t$ using Dijkstra's algorithm restricted to $t$'s biconnected component.
The final path is the concatenation of all three path parts.
